External Prior Guided Internal Prior Learning for Real Noisy Image Denoising

### 2016-11-07 CVPR2017_Review_Unified.tex
add

%%%%%%%%% TITLE
\title{External Patch Group Prior Guided Internal Prior Learning for Real Image Denoising}

@@

Assume we have assigned all internal noisy PGs $\{\mathbf{\overline{Y}}_{n}\}_{n=1}^{N}$ to their corresponding most suitable Gaussians or subspaces in $\{\mathcal{N}(\mathbf{0},\mathbf{\Sigma}_{k})\}_{k=1}^{K}$. For the $k$-th subspace, assume the noisy PGs assigned to it are $\{\mathbf{\overline{Y}}_{k_{n}}\}_{n=1}^{N_{k}}$ such that $\mathbf{\overline{Y}}_{k_{n}}=[\mathbf{\overline{y}}_{k_{n},1},...,\mathbf{\overline{y}}_{k_{n},M}]$ and $\sum_{k=1}^{K}N_{k}=N$. We utilize the external orthogonal dictionary $\mathbf{U}_{k}$ (Equ. (\ref{equ3})) to guide the learning of an orthogonal dictionary for adaptively characterizing the internal PGs in the $k$-th subspace. The learned dictionary defined as $\mathbf{D}_{k}\triangleq[\mathbf{D}_{k,e}\ \mathbf{D}_{k,i}]\in \mathbb{R}^{3p^2\times 3p^2}$ has two parts: the external part $\mathbf{D}_{k,e}=\mathbf{U}_{k}(:,1:3p^2-r)\in\mathbb{R}^{3p^2\times (3p^2-r)}$ is directly obtained from the external dictionary $\mathbf{U}_{k}$, and the internal part $\mathbf{D}_{k,i}$ is consisted of dictionary atoms adaptively learned from the internal noisy PGs $\{\mathbf{\overline{Y}}_{k_{n}}\}_{n=1}^{N_{k}}$. For notation simplicity, we ignore the subspace index $k$ and denote the noisy PGs assigned to each subspace as $\mathbf{Y}\triangleq\{\mathbf{\overline{Y}}_{n}\}_{n=1}^{N} = [\mathbf{\overline{y}}_{1,1},...,\mathbf{\overline{y}}_{1,M},...,\mathbf{\overline{y}}_{N,1},...,\mathbf{\overline{y}}_{N,M}]$. The learning is performed under the weighted sparse coding framework proposed as follows:
\vspace{-2mm}
\begin{equation}\label{equ5}\vspace{-3mm}
\begin{split}
&\min_{\mathbf{D}_{i},\{\bm{\alpha}_{n,m}\}}
\sum_{n=1}^{N}\sum_{m=1}^{M}(\|\mathbf{\overline{y}}_{n,m}-\mathbf{D}\bm{\alpha}_{n,m}\|_{2}^{2}+\sum_{j=1}^{3p^{2}}\lambda_{j}|\bm{\alpha}_{n,m,j}|)
\\
&
s.t.
\quad
\mathbf{D}=[\mathbf{D}_{e}\ \mathbf{D}_{i}],\ \mathbf{D}_{i}^{\top}\mathbf{D}_{i} = \mathbf{I}_{r},\ \mathbf{D}_{e}^{\top}\mathbf{D}_{i} = \mathbf{0},
\end{split}
\end{equation}
where $\bm{\alpha}_{n,m}$ is the sparse coefficient vector of the $m$-th patch $\mathbf{\overline{y}}_{n,m}$ in the $n$-th PG $\mathbf{\overline{Y}}_{n}$ and $\bm{\alpha}_{n,m,j}$ is the $j$-th element of $\bm{\alpha}_{n,m}$. $\lambda_{j}$ is the $j$-th regularization parameter defined as
\vspace{-1mm}
\begin{equation}\label{equ6}\vspace{-1mm}
\lambda_{j} = \lambda/(\sqrt{\mathbf{S}_{j}}+\varepsilon).
\end{equation}
We employ square roots of the singular values in $\mathbf{S}$ (please refer to Equ. (\ref{equ3})) as external prior weights and add a small positive number $\varepsilon$ to avoid zero denominator. Noted that $\mathbf{D}_{e}=\emptyset$ if $r=3p^{2}$ and $\mathbf{D}_{e}=\mathbf{U}_{k}$ if $r=0$. The dictionary $\mathbf{D} = [\mathbf{D}_{e}\ \mathbf{D}_{i}]$ is orthogonal by checking that:
\vspace{-2mm}
\begin{equation}\label{equ7}\vspace{-1mm}
\mathbf{D}^{\top}\mathbf{D} = 
\left[\begin{array}{c}
\mathbf{D}_{e}^{\top}
\\
\mathbf{D}_{i}^{\top}
\end{array}\right]
[\mathbf{D}_{e}\ \mathbf{D}_{i}]
=
\left[\begin{array}{cc}
\mathbf{D}_{e}^{\top}\mathbf{D}_{e}\ \mathbf{D}_{e}^{\top}\mathbf{D}_{i}
\\
\mathbf{D}_{i}^{\top}\mathbf{D}_{e}\ \mathbf{D}_{i}^{\top}\mathbf{D}_{i}
\end{array}\right]
=
\mathbf{I}
\end{equation}

@@

Here we take a detailed analysis on the guidance of the external patch group (PG) prior for the internal noisy PGs of given real noisy images. The guidance comes from at least three aspects: 1) the external prior guides the internal PGs to be clustered into suitable subspaces through MAP in Equ. (\ref{equ4}). The guided subspace selection is more efficient than directly clustering the internal noisy PGs via k-means or Gaussian Mixture Model (GMM). The reason is, by guidance we only need calculate the probabilities via Equ. (\ref{equ4}) for all noisy PGs while by internal clustering via GMM we have to perform time-consuming EM algorithm \cite{em}; 2) the external clean dictionary guides the learning of orthogonal dictionaries more adaptive for internal noisy images. The learning process is very efficient because of closed-form solutions. Besides, the learned orthogonal dictionary also makes the denoising process very efficient under weighted sparse coding framework. 3) the singular values learned by SVD in Equ. (\ref{equ3}) reflect the prior weights of the dictionary atoms and can be used as adaptive parameters for real image denoising.

@@

We evaluate the performance of the proposed framework on denoising real noisy images. The denoising is simultaneously done with the guided internal dictionary learning process. We ignore the index $k\in\{1,...,K\}$ of subspace for notation simplicity. In the denoising stage, for each subspace, the group mean vectors $\{\bm{\mu}_{n}\}_{n=1}^{N}$ of corresponding mean subtracted noisy PGs $\{\mathbf{\overline{Y}}_{n}\}_{n=1}^{N}$ are saved for reconstruction. Until now, we obtain the solutions of sparse coefficient vectors $\{\hat{\bm{\alpha}}_{n,m}^{(T-1)}\}$ in (\ref{equ9}) for $n=1,...,N;m=1,...,M$ and the orthogonal dictionary $\mathbf{D}_{(T)} = [\mathbf{D}_{e}\ \mathbf{D}_{i}^{(T)}]$ in Equ. (\ref{equ10}). Then the $m$-th latent clean patch $\hat{\mathbf{y}}_{n,m}$ in the $n$-th PG $\mathbf{Y}_{n}$ is recovered by 
\vspace{-2mm}
\begin{equation}\label{equ12}\vspace{-2mm}
\hat{\mathbf{y}}_{n,m}=\mathbf{D}_{(T)}\hat{\bm{\alpha}}_{n,m}+\bm{\mu}_{n},
\end{equation}
where $n=1,...,N;m=1,...,M$.
\begin{table}\label{alg1}
\begin{tabular}{l}
\hline
\textbf{Alg. 1}: External Patch Group (PG) Prior Guided Internal
\\
\quad \quad \quad PG Prior Learning for Image Denoising
\\
\hline
\textbf{Input:} Noisy image $\mathbf{y}$, external PG prior GMM model
\\
\textbf{Output:} The denoised image $\hat{\mathbf{x}}$.
\\
\textbf{Initialization:} $\hat{\mathbf{x}}^{(0)}=\mathbf{y}$;
\\
\textbf{for} $Ite = 1:IteNum$ \textbf{do}
\\
1. Extracting internal PGs from $\hat{\mathbf{x}}^{(Ite-1)}$;
\\
%\textbf{Guided Internal Subspace Selection:}
%\\
\quad\textbf{for} each PG $\mathbf{Y}_{n}$ \textbf{do}
\\
2.\quad Calculate group mean vector $\boldsymbol{\upmu}_{n}$ and form 
\\
\quad \ \ \ mean subtracted PG $\mathbf{\overline{Y}}_{n}$;
\\
3.\quad Subspace selection via Equ. (\ref{equ4});
\\
\quad\textbf{end for}
\\
%\textbf{Guided Internal Orthogonal Dictionary Learning:}
%\\
\quad\textbf{for} the PGs in each Subspace \textbf{do}
\\
4.\quad External PG prior Guided Internal Orthogonal
\\
\quad \ \ \ Dictionary Learning by solving (\ref{equ5});
\\
5.\quad Recover each patch in all PGs via Equ. (\ref{equ12});
\\
\quad\textbf{end for}
\\
6. Aggregate the recovered PGs of all subspaces to form
\\
\quad the recovered image $\hat{\mathbf{x}}^{(Ite)}$;
\\
\textbf{end for}
\\
\hline
\end{tabular}
\end{table}
The latent clean image $\hat{\mathbf{x}}$ is reconstructed by aggregating all the estimated PGs. Similar to \cite{pgpd}, we perform the above denoising procedures for several iterations for better denoising outputs. The proposed denoising algorithm is summarized in Alg. 1.

@@

\subsection{Comparison with External and Internal Priors}
In this section, we compare our proposed method on real image denoising with external prior based method (denoted as ``External") and internal prior based method (denoted as ``Internal"). The three methods perform denoising under the same framework as Equ. (\ref{equ5}). The ``External" method employs the external dictionaries (i.e., $r=0$ in Equ. (\ref{equ5})) for denoising, while the ``Internal" method employ the GMM (also with $K=32$ Gaussians) to directly cluster the internal noisy PGs into multiple subspaces, and for each subspace perform denoising with fully learned dictionary (i.e., $r=3p^{2}$ in Equ. (\ref{equ5})). All parameters of the three methods are tuned to achieve best performance.

>>>>

%%%%%%%%% TITLE
\title{External Prior Guided Internal Prior Learning for Real Noisy Image Denoising}

@@

Assume we have assigned all internal noisy PGs $\{\mathbf{\overline{Y}}_{n}\}_{n=1}^{N}$ to their corresponding most suitable Gaussians or subspaces in $\{\mathcal{N}(\mathbf{0},\mathbf{\Sigma}_{k})\}_{k=1}^{K}$. For the $k$-th subspace, assume the noisy PGs assigned to it are $\{\mathbf{\overline{Y}}_{k_{n}}\}_{n=1}^{N_{k}}$ such that $\mathbf{\overline{Y}}_{k_{n}}=[\mathbf{\overline{y}}_{k_{n},1},...,\mathbf{\overline{y}}_{k_{n},M}]$ and $\sum_{k=1}^{K}N_{k}=N$. We utilize the external orthogonal dictionary $\mathbf{U}_{k}$ (Eqn. (\ref{equ3})) to guide the learning of an orthogonal dictionary for adaptively characterizing the internal PGs in the $k$-th subspace. The learned dictionary defined as $\mathbf{D}_{k}\triangleq[\mathbf{D}_{k,e}\ \mathbf{D}_{k,i}]\in \mathbb{R}^{3p^2\times 3p^2}$ has two parts: the external part $\mathbf{D}_{k,e}=\mathbf{U}_{k}(:,1:3p^2-r)\in\mathbb{R}^{3p^2\times (3p^2-r)}$ is directly obtained from the external dictionary $\mathbf{U}_{k}$, and the internal part $\mathbf{D}_{k,i}$ is consisted of dictionary atoms adaptively learned from the internal noisy PGs $\{\mathbf{\overline{Y}}_{k_{n}}\}_{n=1}^{N_{k}}$. For notation simplicity, we ignore the subspace index $k$ and denote the noisy PGs assigned to each subspace as $\mathbf{Y}\triangleq\{\mathbf{\overline{Y}}_{n}\}_{n=1}^{N} = [\mathbf{\overline{y}}_{1,1},...,\mathbf{\overline{y}}_{1,M},...,\mathbf{\overline{y}}_{N,1},...,\mathbf{\overline{y}}_{N,M}]$. The learning is performed under the weighted sparse coding framework proposed as follows:
\vspace{-2mm}
\begin{equation}\label{equ5}\vspace{-3mm}
\begin{split}
&\min_{\mathbf{D}_{i},\{\bm{\alpha}_{n,m}\}}
\sum_{n=1}^{N}\sum_{m=1}^{M}(\|\mathbf{\overline{y}}_{n,m}-\mathbf{D}\bm{\alpha}_{n,m}\|_{2}^{2}+\sum_{j=1}^{3p^{2}}\lambda_{j}|\bm{\alpha}_{n,m,j}|)
\\
&
s.t.
\quad
\mathbf{D}=[\mathbf{D}_{e}\ \mathbf{D}_{i}],\ \mathbf{D}_{i}^{\top}\mathbf{D}_{i} = \mathbf{I}_{r},\ \mathbf{D}_{e}^{\top}\mathbf{D}_{i} = \mathbf{0},
\end{split}
\end{equation}
where $\bm{\alpha}_{n,m}$ is the sparse coefficient vector of the $m$-th patch $\mathbf{\overline{y}}_{n,m}$ in the $n$-th PG $\mathbf{\overline{Y}}_{n}$ and $\bm{\alpha}_{n,m,j}$ is the $j$-th element of $\bm{\alpha}_{n,m}$. $\lambda_{j}$ is the $j$-th regularization parameter defined as
\vspace{-1mm}
\begin{equation}\label{equ6}\vspace{-1mm}
\lambda_{j} = \lambda/(\sqrt{\mathbf{S}_{j}}+\varepsilon).
\end{equation}
We employ square roots of the singular values in $\mathbf{S}$ (please refer to Eqn. (\ref{equ3})) as external prior weights and add a small positive number $\varepsilon$ to avoid zero denominator. Noted that $\mathbf{D}_{e}=\emptyset$ if $r=3p^{2}$ and $\mathbf{D}_{e}=\mathbf{U}_{k}$ if $r=0$. The dictionary $\mathbf{D} = [\mathbf{D}_{e}\ \mathbf{D}_{i}]$ is orthogonal by checking that:
\vspace{-2mm}
\begin{equation}\label{equ7}\vspace{-1mm}
\mathbf{D}^{\top}\mathbf{D} = 
\left[\begin{array}{c}
\mathbf{D}_{e}^{\top}
\\
\mathbf{D}_{i}^{\top}
\end{array}\right]
[\mathbf{D}_{e}\ \mathbf{D}_{i}]
=
\left[\begin{array}{cc}
\mathbf{D}_{e}^{\top}\mathbf{D}_{e}\ \mathbf{D}_{e}^{\top}\mathbf{D}_{i}
\\
\mathbf{D}_{i}^{\top}\mathbf{D}_{e}\ \mathbf{D}_{i}^{\top}\mathbf{D}_{i}
\end{array}\right]
=
\mathbf{I}
\end{equation}

@@

Here we take a detailed analysis on the guidance of the external patch group (PG) prior for the internal noisy PGs of given real noisy images. The guidance comes from at least three aspects: 1) the external prior guides the internal PGs to be clustered into suitable subspaces through MAP in Eqn. (\ref{equ4}). The guided subspace selection is more efficient than directly clustering the internal noisy PGs via k-means or Gaussian Mixture Model (GMM). The reason is, by guidance we only need calculate the probabilities via Eqn. (\ref{equ4}) for all noisy PGs while by internal clustering via GMM we have to perform time-consuming EM algorithm \cite{em}; 2) the external clean dictionary guides the learning of orthogonal dictionaries more adaptive for internal noisy images. The learning process is very efficient because of closed-form solutions. Besides, the learned orthogonal dictionary also makes the denoising process very efficient under weighted sparse coding framework. 3) the singular values learned by SVD in Eqn. (\ref{equ3}) reflect the prior weights of the dictionary atoms and can be used as adaptive parameters for real image denoising.

@@

We evaluate the performance of the proposed framework on denoising real noisy images. The denoising is simultaneously done with the guided internal dictionary learning process. We ignore the index $k\in\{1,...,K\}$ of subspace for notation simplicity. In the denoising stage, for each subspace, the group mean vectors $\{\bm{\mu}_{n}\}_{n=1}^{N}$ of corresponding mean subtracted noisy PGs $\{\mathbf{\overline{Y}}_{n}\}_{n=1}^{N}$ are saved for reconstruction. Until now, we obtain the solutions of sparse coefficient vectors $\{\hat{\bm{\alpha}}_{n,m}^{(T-1)}\}$ in (\ref{equ9}) for $n=1,...,N;m=1,...,M$ and the orthogonal dictionary $\mathbf{D}_{(T)} = [\mathbf{D}_{e}\ \mathbf{D}_{i}^{(T)}]$ in Eqn. (\ref{equ10}). Then the $m$-th latent clean patch $\hat{\mathbf{y}}_{n,m}$ in the $n$-th PG $\mathbf{Y}_{n}$ is recovered by 
\vspace{-2mm}
\begin{equation}\label{equ12}\vspace{-2mm}
\hat{\mathbf{y}}_{n,m}=\mathbf{D}_{(T)}\hat{\bm{\alpha}}_{n,m}+\bm{\mu}_{n},
\end{equation}
where $n=1,...,N;m=1,...,M$.
\begin{table}\label{alg1}
\begin{tabular}{l}
\hline
\textbf{Alg. 1}: External Patch Group (PG) Prior Guided Internal
\\
\quad \quad \quad PG Prior Learning for Image Denoising
\\
\hline
\textbf{Input:} Noisy image $\mathbf{y}$, external PG prior GMM model
\\
\textbf{Output:} The denoised image $\hat{\mathbf{x}}$.
\\
\textbf{Initialization:} $\hat{\mathbf{x}}^{(0)}=\mathbf{y}$;
\\
\textbf{for} $Ite = 1:IteNum$ \textbf{do}
\\
1. Extracting internal PGs from $\hat{\mathbf{x}}^{(Ite-1)}$;
\\
%\textbf{Guided Internal Subspace Selection:}
%\\
\quad\textbf{for} each PG $\mathbf{Y}_{n}$ \textbf{do}
\\
2.\quad Calculate group mean vector $\boldsymbol{\upmu}_{n}$ and form 
\\
\quad \ \ \ mean subtracted PG $\mathbf{\overline{Y}}_{n}$;
\\
3.\quad Subspace selection via Eqn. (\ref{equ4});
\\
\quad\textbf{end for}
\\
%\textbf{Guided Internal Orthogonal Dictionary Learning:}
%\\
\quad\textbf{for} the PGs in each Subspace \textbf{do}
\\
4.\quad External PG prior Guided Internal Orthogonal
\\
\quad \ \ \ Dictionary Learning by solving (\ref{equ5});
\\
5.\quad Recover each patch in all PGs via Eqn. (\ref{equ12});
\\
\quad\textbf{end for}
\\
6. Aggregate the recovered PGs of all subspaces to form
\\
\quad the recovered image $\hat{\mathbf{x}}^{(Ite)}$;
\\
\textbf{end for}
\\
\hline
\end{tabular}
\end{table}
The latent clean image $\hat{\mathbf{x}}$ is reconstructed by aggregating all the estimated PGs. Similar to \cite{pgpd}, we perform the above denoising procedures for several iterations for better denoising outputs. The proposed denoising algorithm is summarized in Alg. 1.

@@

\subsection{Comparison with External and Internal Priors}
In this section, we compare our proposed method on real image denoising with external prior based method (denoted as ``External") and internal prior based method (denoted as ``Internal"). The three methods perform denoising under the same framework as Eqn. (\ref{equ5}). The ``External" method employs the external dictionaries (i.e., $r=0$ in Eqn. (\ref{equ5})) for denoising, while the ``Internal" method employ the GMM (also with $K=32$ Gaussians) to directly cluster the internal noisy PGs into multiple subspaces, and for each subspace perform denoising with fully learned dictionary (i.e., $r=3p^{2}$ in Eqn. (\ref{equ5})). All parameters of the three methods are tuned to achieve best performance.
